---
title: "Enrichment"
output:
  html_document:
    pairs.df_print: paged
---


```{r eurofiles, cache=TRUE}
library(tidyr)
library(dplyr)
library(data.table)
library(ggplot2)

source('./enrichment.R')

rootdir <- '/scratch/groups/abattle4/victor/WatershedAFR/'
datadir <- paste0(rootdir,'/','data/')

```
##European Samples

```{r europrep, cache=TRUE, eval=FALSE}
# files needed for european variants
euro_z_scores = '/scratch/groups/abattle4/victor/WatershedAFR/data/outlier_calling/gtexV8.outlier.controls.v8ciseQTLs.globalOutliers.removed.medz.txt'
 # table of gene-individual pairs with their rare variants
euro_rare_var_pairs =
  # '/scratch/groups/abattle4/victor/WatershedAFR/data/rare_variants/old/gene-EUR-rv.txt' # table of gene-individual pairs with their rare variants
'/scratch/groups/abattle4/victor/WatershedAFR/data/rare_variants/gene-EUR-rv.txt'

euro_pop_list = '/scratch/groups/abattle4/victor/WatershedAFR/data/data_prep/gtex_v8_wgs_individuals_EUR.txt'  #list of samples from the desired population

```


Here are the results of enrichment on the European samples using thresholds 1, 1.2, 1.4, 1.5, 1.6, 1.8, 2, 2.5, 3, 3.5, 4, 4.5, 5:
```{r euroretrieve, cache=TRUE}

euro.output.3 <- readRDS(paste0(datadir,'/european_1to5.rds'))
euro.output.3
```


###Binning Thresholds

#### Nothing above threshold + 0.5
```{r eurobin, cache=TRUE}

euro.output4 <- readRDS(paste0(datadir,'/european_1to5_binw05.rds'))
euro.output4

```

#### Cutoff at threshold + 1
```{r eurobin1, cache=TRUE}
euro.output5 <- readRDS(paste0(datadir,'/european_1to5_binw1.rds'))
euro.output5
```



## African American Samples
```{r afroprep}
# files needed for african variants
afr_z_scores_file = '/scratch/groups/abattle4/victor/WatershedAFR/data/outlier_calling/gtexV8.outlier.controls.v8ciseQTLs.globalOutliers.removed.medz.txt'
afr_rare_var_pairs_file = '/scratch/groups/abattle4/victor/WatershedAFR/data/rare_variants/gene-AFR-rv.txt' # table of gene-individual pairs with their rare variants
afr_pop_list_file = '/scratch/groups/abattle4/victor/WatershedAFR/data/data_prep/gtex_v8_wgs_individuals_AFR.txt'
```

Here are the results of enrichment on the African samples using thresholds 1, 1.2, 1.4, 1.5, 1.6, 1.8, 2, 2.5, 3, 3.5, 4, 4.5, 5:

```{r afrretrieve, cache=TRUE}

afr.output.3 <- readRDS(paste0(datadir,'/african_1to5.rds'))
afr.output.3$enrichment$binw=0
afr.output.3

```


###Binning Thresholds
#### Cutoff at threshold + 0.5
```{r afrbin05_retrieve, cache=TRUE}
afr.output.4 <- readRDS(paste0(datadir,'/african_1to5_binw05.rds'))
afr.output.4$enrichment$binw=0.5
afr.output.4$plot2 <- afr.output.4$plot + ggtitle("Enrichment of Rare Variants in Outlier Genes -- African (bw=0.5)")
afr.output.4

```

#### Cutoff at threshold + 1
```{r afrbin05_retrieve, cache=TRUE}
afr.output.5 <- readRDS(paste0(datadir,'/african_1to5_binw1.rds'))
afr.output.5$enrichment$binw=1
afr.output.5$plot <-afr.output.5$plot + ggtitle("Enrichment of Rare Variants in Outlier Genes -- African (bw=1)")
afr.output.5

```


#### Cutoff at threshold + 0.25
```{r afrbin025_retrieve, cache=TRUE}
afr.output.6 <- readRDS(paste0(datadir,'/african_1to5_binw025.rds'))
afr.output.6$enrichment$binw=0.25
afr.output.6$plot <- afr.output.6$plot + ggtitle("Enrichment of Rare Variants in Outlier Genes -- African (bw=0.25)")
afr.output.6

```

#### Cutoff at threshold + 0.15
```{r afrbin015_retrieve, cache=TRUE}
afr.output.7 <- readRDS(paste0(datadir,'/african_1to5_binw015.rds'))
afr.output.7$enrichment$binw=0.15
afr.output.7$plot <- afr.output.7$plot + ggtitle("Enrichment of Rare Variants in Outlier Genes -- African (bw=0.15)")
afr.output.7

```



### AFR Enrichment Summary Graphs

```{r sumg1, cache = TRUE}
afr.df <- bind_rows(afr.output.3$enrichment, afr.output.4$enrichment, afr.output.5$enrichment, afr.output.6$enrichment, afr.output.7$enrichment) %>%
  mutate(binw=factor(binw))

afr.df %>%
  ggplot(aes(x = sortcol, y = enrichment, col=binw)) +
  geom_line(size = .5) + 
    #geom_text(aes(label = outlier_all_pairs), hjust = 0.7, vjust = -4) + 
    geom_errorbar(aes(ymax = upper_CI, ymin = lower_CI)) + 
  labs(title="Enrichment of rare variants in outlier genes (AFR)", x = "Z-score threshold")

afr.df %>% select(sortcol,binw,inlier_all_pairs, outlier_all_pairs,outofbounds_all_pairs, thresh) %>% pivot_longer(!c(sortcol,thresh,binw), values_drop_na=TRUE, names_to="group", values_to="count") %>%
ggplot(aes(x=binw, fill=group)) +
  geom_col(aes(y=count),position="dodge") +
  facet_wrap(vars(thresh)) +
  scale_y_log10() +
  ggtitle("All Pairs")


afr.df %>% select(sortcol,binw,inlier_rare_pairs, outlier_rare_pairs,outofbounds_rare_pairs, thresh) %>% pivot_longer(!c(sortcol,thresh,binw), values_drop_na=TRUE, names_to="group", values_to="count") %>%
ggplot(aes(x=binw, fill=group)) +
  geom_col(aes(y=count),position="dodge") +
  facet_wrap(vars(thresh)) +
  scale_y_log10() +
  ggtitle("Rare Pairs")
```





## Watershed 
```{r wtrshdprep}
wtrshd_z_scores_file = '/scratch/groups/abattle4/victor/WatershedAFR/data/outlier_calling/gtexV8.outlier.controls.v8ciseQTLs.globalOutliers.removed.medz.txt'
wtrshd_rare_var_pairs_file = '/work-zfs/abattle4/bstrober/rare_variant/gtex_v8/splicing/input_data/variant_calls/all_rare_variants_SNPs_10kb_genebody.txt' # table of gene-individual pairs with their rare variants
wtrshd_pop_list_file = '/scratch/groups/abattle4/victor/WatershedAFR/data/data_prep/gtex_v8_wgs_individuals_EUR.txt'

```


```{r wtrshdenrich, cache=TRUE}
wtrshd_output_prefix <- 'gtex_v8_wgs_individuals_wtshd'

rare.colnames <- c("Ind", "Gene", "Chrom", "Start", "End")

wtrshd.output <- enrichmentTipTail(pop_subset_file = wtrshd_pop_list_file, z_scores_file = wtrshd_z_scores_file, rare_var_pairs_file = wtrshd_rare_var_pairs_file, output_dir = datadir, title.mod.pop = "Watershed European", output_prefix = wtrshd_output_prefix, col.names = rare.colnames)


wtrshd.output


```

Results of enrichment on watershed output using thresholds 1, 1.2, 1.4, 1.5, 1.6, 1.8, 2, 2.5, 3, 3.5, 4, 4.5, 5:
```{r wtrshdretrieve, cache=TRUE}

wtrshd.output.3 <- readRDS(paste0(datadir,'/wtrshdeuro_1to5.rds'))
wtrshd.output.3

```

```{r notes}
# pdf(paste0('/scratch/groups/abattle4/victor/WatershedAFR/data/figures','/enrichment_eur.pdf'))
```
