# Visualize Rare Variants and Outliers

```{r}
library(data.table)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggVennDiagram)


# functions

# rv.count.df - dataframe with count of rare variants per gene
# heading - heading for histogram title
plot_hist = function(rv.count, heading) {
  
  # plot histogram
  title = paste0("Rare variants per gene - ", heading)
  p = ggplot(rv.count, aes(x=n)) + 
    geom_histogram(bins = 200) +
    labs(title="Rare variants per gene", x = "# of rare variants", y = "# of genes")
  print(p)
}

# rare variant files
rv_ws_file = '/work-zfs/abattle4/bstrober/rare_variant/gtex_v8/splicing/input_data/variant_calls/all_rare_variants_SNPs_10kb_genebody.txt'
rv_eur_file = '/scratch/groups/abattle4/victor/WatershedAFR/data/rare_variants/gene-EUR-rv.txt'
rv_afr_file = '/scratch/groups/abattle4/victor/WatershedAFR/data/rare_variants/gene-AFR-rv.txt'
```

```{r}
# Watershed 

# open file
rv_ws = read.table(rv_ws_file, header = FALSE, check.names = FALSE)
colnames(rv_ws) = c("Ind", "Gene", "Chrom", "Start", "End")

# count rare variants per gene
rv_ws.count = rv_ws %>% count(Gene)

plot_hist(rv_ws.count, heading = "Watershed")

```

```{r}
# European 

# open files
rv_eur = read.table(rv_eur_file, header = TRUE, check.names = FALSE)

# count rare variants per gene
rv_eur.count = rv_eur %>% count(Gene)

plot_hist(rv_eur.count, heading = "European")
```

```{r}
# African 

# open files
rv_afr = read.table(rv_afr_file, header = TRUE, check.names = FALSE)

# count rare variants per gene
rv_afr.count = rv_afr %>% count(Gene)

plot_hist(rv_afr.count, heading = "Watershed")
```

```{r}
# combined histogram
rv_ws.count = rv_ws.count %>% mutate(source = "Watershed")
rv_eur.count = rv_eur.count %>% mutate(source = "European")
rv_afr.count = rv_afr.count %>% mutate(source = "African")

rv.count = bind_rows(rv_ws.count, rv_eur.count, rv_afr.count)
rv.count = rv.count %>% arrange(desc(n))

ggplot(rv.count, aes(x=n, fill=source)) + geom_histogram(alpha=0.2, bins = 100) +
  labs(title="Rare variants per gene", x = "# of rare variants", y = "# of genes")

```

```{r}
# split into less than and greater than percentile
percentile = 0.95
thresh = quantile(rv.count$n, percentile)

rv.count.lower = rv.count %>% filter(n <= thresh)
rv.count.upper = rv.count %>% filter(n > thresh)

hist.lower = ggplot(rv.count.lower, aes(x=n, fill=source)) + geom_histogram(alpha=0.2, binwidth = 100) +
  labs(title=paste0("Rare variants per gene (<=", percentile, ")"), x = "# of rare variants", y = "# of genes")

hist.upper = ggplot(rv.count.upper, aes(x=n, fill=source)) + geom_histogram(alpha=0.2) +
  labs(title=paste0("Rare variants per gene (>", percentile, ")"), x = "# of rare variants", y = "# of genes")

print(hist.lower)
print(hist.upper)

```

```{r}
# Genes with top percentile of rare variants within each source
thresh_ws = quantile(rv_ws.count$n, percentile)
thresh_eur = quantile(rv_eur.count$n, percentile)
thresh_afr = quantile(rv_afr.count$n, percentile)

rv_ws.count.upper = rv_ws.count %>% filter(n > thresh_ws) %>% arrange(desc(n))
rv_eur.count.upper = rv_eur.count %>% filter(n > thresh_eur) %>% arrange(desc(n))
rv_afr.count.upper = rv_afr.count %>% filter(n > thresh_afr) %>% arrange(desc(n))

top_genes_ws = unique(rv_ws.count.upper$Gene)
top_genes_eur = unique(rv_eur.count.upper$Gene)
top_genes_afr = unique(rv_afr.count.upper$Gene)
